scrubr
======

```{r echo=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

[![R-check](https://github.com/ropensci/scrubr/workflows/R-check/badge.svg)](https://github.com/ropensci/scrubr/actions?query=workflow%3AR-check)
[![cran checks](https://cranchecks.info/badges/worst/scrubr)](https://cranchecks.info/pkgs/scrubr)
[![codecov.io](https://codecov.io/github/ropensci/scrubr/coverage.svg?branch=master)](https://codecov.io/github/ropensci/scrubr?branch=master)
[![rstudio mirror downloads](https://cranlogs.r-pkg.org/badges/scrubr?color=ff69b4)](https://github.com/r-hub/cranlogs.app)
[![cran version](https://www.r-pkg.org/badges/version/scrubr)](https://cran.r-project.org/package=scrubr)

__Clean Biological Occurrence Records__

Clean using the following use cases (checkmarks indicate fxns exist - not necessarily complete):

- [x] Impossible lat/long values: e.g., latitude 75
- [x] Incomplete cases: one or the other of lat/long missing
- [x] Unlikely lat/long values: e.g., points at 0,0
- [x] Deduplication: try to identify duplicates, esp. when pulling data from multiple sources, e.g., can try to use occurrence IDs, if provided
- [x] Date based cleaning
* [x] Outside political boundary: User input to check for points in the wrong country, or points outside of a known country
* [x] Taxonomic name based cleaning: via `taxize` (one method so far)
* Political centroids: unlikely that occurrences fall exactly on these points, more likely a
default position (Draft function started, but not exported, and commented out). see [issue #6](https://github.com/ropensci/scrubr/issues/6)
* Herbaria/Museums: many specimens may have location of the collection they are housed in, see [issue #20](https://github.com/ropensci/scrubr/issues/20)
* Habitat type filtering: e.g., fish should not be on land; marine fish should not be in fresh water
* Check for contextually wrong values: That is, if 99 out of 100 lat/long coordinates are within the continental US, but 1 is in China, then perhaps something is wrong with that one point
* Collector/recorder names: see [issue #19](https://github.com/ropensci/scrubr/issues/19)
* ...

A note about examples: We think that using a piping workflow with `%>%` makes code easier to
build up, and easier to understand. However, in some examples we provide examples without the pipe
to demonstrate traditional usage.

## Install

Stable CRAN version

```{r eval=FALSE}
install.packages("scrubr")
```

Development version

```{r eval=FALSE}
remotes::install_github("ropensci/scrubr")
```

```{r}
library("scrubr")
```

## Coordinate based cleaning

```{r}
data("sampledata1")
```

Remove impossible coordinates (using sample data included in the pkg)

```{r}
# coord_impossible(dframe(sample_data_1)) # w/o pipe
dframe(sample_data_1) %>% coord_impossible()
```

Remove incomplete coordinates

```{r}
# coord_incomplete(dframe(sample_data_1)) # w/o pipe
dframe(sample_data_1) %>% coord_incomplete()
```

Remove unlikely coordinates (e.g., those at 0,0)

```{r}
# coord_unlikely(dframe(sample_data_1)) # w/o pipe
dframe(sample_data_1) %>% coord_unlikely()
```

Do all three

```{r}
dframe(sample_data_1) %>%
  coord_impossible() %>%
  coord_incomplete() %>%
  coord_unlikely()
```

Don't drop bad data

```{r}
dframe(sample_data_1) %>% coord_incomplete(drop = TRUE) %>% NROW
dframe(sample_data_1) %>% coord_incomplete(drop = FALSE) %>% NROW
```


## Deduplicate

```{r}
smalldf <- sample_data_1[1:20, ]
# create a duplicate record
smalldf <- rbind(smalldf, smalldf[10,])
row.names(smalldf) <- NULL
# make it slightly different
smalldf[21, "key"] <- 1088954555
NROW(smalldf)
dp <- dframe(smalldf) %>% dedup()
NROW(dp)
attr(dp, "dups")
```

## Dates

Standardize/convert dates

```{r}
df <- sample_data_1
# date_standardize(dframe(df), "%d%b%Y") # w/o pipe
dframe(df) %>% date_standardize("%d%b%Y")
```

Drop records without dates

```{r}
NROW(df)
NROW(dframe(df) %>% date_missing())
```

Create date field from other fields

```{r}
dframe(sample_data_2) %>% date_create(year, month, day)
```

## Ecoregion

Filter by FAO areas

```{r eval=FALSE}
wkt <- 'POLYGON((72.2 38.5,-173.6 38.5,-173.6 -41.5,72.2 -41.5,72.2 38.5))'
manta_ray <- rgbif::name_backbone("Mobula alfredi")$usageKey
res <- rgbif::occ_data(manta_ray, geometry = wkt, limit=300, hasCoordinate = TRUE)
dat <- sf::st_as_sf(res$data, coords = c("decimalLongitude", "decimalLatitude"))
dat <- sf::st_set_crs(dat, 4326)
mapview::mapview(dat)
tmp <- eco_region(dframe(res$data), dataset = "fao", region = "OCEAN:Indian")
tmp <- tmp[!is.na(tmp$decimalLongitude), ]
tmp2 <- sf::st_as_sf(tmp, coords = c("decimalLongitude", "decimalLatitude"))
tmp2 <- sf::st_set_crs(tmp2, 4326)
mapview::mapview(tmp2)
```

## Meta

* Please [report any issues or bugs](https://github.com/ropensci/scrubr/issues).
* License: MIT
* Get citation information for `scrubr` in R doing `citation(package = 'scrubr')`
* Please note that this package is released with a [Contributor Code of Conduct](https://ropensci.org/code-of-conduct/). By contributing to this project, you agree to abide by its terms.
