cleanoccs
=========

```{r echo=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

[![Build Status](https://travis-ci.org/sckott/cleanoccs.svg?branch=master)](https://travis-ci.org/sckott/cleanoccs)
[![codecov.io](http://codecov.io/github/sckott/cleanoccs/coverage.svg?branch=master)](http://codecov.io/github/sckott/cleanoccs?branch=master)

__Clean Biological Occurrence Records__

Clean using the following use cases:

* Impossible lat/long values: e.g., latitude 75
* Incomplete cases: one or the other of lat/long missing
* Unlikely lat/long values: e.g., points at 0,0
* Political centroids: unlikely that occurrences fall exactly on these points, more likely a
default position
* Herbaria/Museums: many specimens may have location of the collection they are housed in
* Habitat type filtering: e.g., fish should not be on land; marine fish should not be in fresh water
* Deduplication: try to identify duplicates, esp. when pulling data from multiple sources, e.g., can try to use occurrence IDs, if provided
* Check for contextually wrong values: That is, if 99 out of 100 lat/long coordinates are within the continental US, but 1 is in China, then perhaps something is wrong with that one point
* Outside political boundary: User input to check for points in the wrong country, or points outside of a known country
* Taxonomic name based cleaning: via `taxize`
* Date based cleaning perhaps?
* ...

## Install

```{r eval=FALSE}
devtools::install_github("sckott/cleanoccs")
```

```{r}
library("cleanoccs")
```

## Coordinate based cleaning

Remove impossible coordinates (using sample data included in the pkg)

```{r}
clean_df(sample_data_1) %>% coord_impossible()
```

Remove impossible coordinates

```{r}
clean_df(sample_data_1) %>% coord_incomplete()
```

Remove unlikely points (e.g., those at 0,0)

```{r}
clean_df(sample_data_1) %>% coord_unlikely()
```

Do all three

```{r}
clean_df(sample_data_1) %>%
  coord_impossible() %>%
  coord_incomplete() %>%
  coord_unlikely()
```

Don't drop bad data

```{r}
clean_df(sample_data_1) %>% coord_incomplete(drop = TRUE) %>% NROW
clean_df(sample_data_1) %>% coord_incomplete(drop = FALSE) %>% NROW
```


## Dedpulicate

```{r}
smalldf <- sample_data_1[1:20, ]
# create a duplicate record
smalldf <- rbind(smalldf, smalldf[10,])
# make it slightly different
smalldf[21, "key"] <- 1088954555 
NROW(smalldf)
dp <- clean_df(smalldf) %>% dedup()
NROW(dp)
attr(dp, "dups")
```

## Meta

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
